Scheduling and Vectorization for MPC

### 2022-03-05 paper_SIMD.tex
Commented out Longest Common SuperSeq stuff to test whether latex github action works.
Commented out section to have paper compile on Github.
Added first draft of correctness argument section.

@@

@@

\ana{Ishaq?}

\ishaq{The follow text is all from Linsay's thesis}.
\brevised{
We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, 
but as an example, $A$ stands for the ADD MPC instruction 
and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or
an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
between classical scheduling, as studied in parallel computing, and MPC scheduling.

@@

To see the reduction, 
suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
$S_i$, i.e., $P'$ is a common supersequence 
of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
which is a contradiction since $P$ is a schedule with minimal cost.)
}

@@

\ana{Add the AI-based proof here.}

>>>>

\ana{Commenting out this section to have paper compile on Github. Will revert!}
\begin{comment}

@@

\end{comment}

@@

\ana{Commented out this section to have paper compile on Github. Will revert!}

\begin{comment}
\ishaq{The follow text is all from Linsay's thesis}.
\brevised{
We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, 
but as an example, $A$ stands for the ADD MPC instruction 
and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or
an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
between classical scheduling, as studied in parallel computing, and MPC scheduling.

@@

To see the reduction, 
suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
$S_i$, i.e., $P'$ is a common supersequence 
of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
which is a contradiction since $P$ is a schedule with minimal cost.)
}
\end{comment}

@@

We build a correctness argument that loosely follows the theory of Abstract Interpretation. 
We define the syntax of MPC Source programs. The domain of MPC Source programs
expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$. 
We then define the \emph{linearization} of an MPC Source program as an interpretation 
over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$. 
Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$ 
in terms of def-use relations. We define a partial order over elements of $C$ as well,
in terms of def-use relations in the concrete domain $C$. We prove two theorems
that state (informally) that the schedule corresponding to the original program computes the same 
result as the schedule corresponding to the vectorized program.

\begin{figure}%[!b]
%\small
\[
\begin{array}{l@{~~~~~}l}
  \begin{array}{l@{~}l@{~~~}l}
  s &::= s;s & sequence \\
     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\ 
     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
     & \mid \x = \y & \mathit{propagation} \\
     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
  \end{array}    \\
\end{array}
\]
\vspace{-3mm}
\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
\vspace{-5mm}
\end{figure}

\paragraph{MPC Source Syntax}

\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source. 
There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array 
$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below. 
Thus, $i$ and $k$ are in scope.
\begin{verbatim}
FOR 0 <= i < I:
    ...
    FOR 0 <= k < K:
        x[i,J,k] = ...
\end{verbatim}

Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements, 
and $\mathit{sequence, loop}$ are \emph{compound statements}.

\paragraph{Linearization}

Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
MPC Source syntax, as it is standard. The concretization of each one 
of the base statements is as follows: 
\[\begin{array}{l}
\gamma(\x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k]) = \\ 
\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
\end{array}\]
meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.

The concretization of the FOR statement is as follows:
\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] ; \gamma(s)[1/i] ; ... \gamma(s)[I-1/i] \]
$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.

\paragraph{Partial Orders}

For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
 $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.

 Let $a_0, a_1$ be two MPC Source programs in $A$. 
 Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on 
 the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case. 
 Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.

 \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
 $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
 \end{definition}

 The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
 to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
 indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
 E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.

\paragraph{Theorems}

\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
\end{theorem}

\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm. 
We have that $a_0 \le a_1$. 
\end{theorem}

\ana{Write the proof sketch and final argument, etc.}
